r5r: rapid realistic routing on multimodal transport networks with R5 in R

### 2021-01-18 r-package/paper2020/paper2020.Rmd
paper edits, plus fix stepniak citation

<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) ... 666666666666

@@

The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either city or country-level analysis. It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM), and it  automatically combines public transport data if they are provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format .

@@

-   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>. This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS datasets to build a routable network object.

@@

Both routing functions are versatile so users can easily set customized inputs such as transport modes, departure date and times, walking and cycling speeds, maximum trip duration, walking distances and number of public transport transfers. We will focus on results obtained from `travel_time_matrix()` on the findings discussed in the following section.

@@

After installed with the `install.packages("r5r")` command, the package can be attached (alongside some other packages used in this article), as follows:

@@

### Build a routable transport network

To build a routable transport network with r5r and load it in memory, the user needs one command line pointing to the location on disk where the input data is stored. In the examples of this paper, we will be using the provided Porto Alegre dataset.

@@

### Calculate a travel time matrix

The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as *maximum trip duration* or *walking distance*, and outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.

To account for variations on service levels across the day [@stepniak2019impact], the user can use the `time_window` parameter. In this case, several matrices are calculated usinh every minute within the set time window as a departure time. By default the function outputs the 50th percentile (i.e. the median) travel time between each origin and destination pair within that time window. The `percentiles` parameter allows the user to retrieve travel time estimates at different points of the distribution. An example of the `travel_time_matrix()` function's usage is presented below:

<!-- However, this can provide biased estimates due to how service levels varies across the day (reference to https://www.sciencedirect.com/science/article/pii/S0966692318305209).  -->

<!-- To account for these variations,  -->

>>>>

<!-- More importantly, those tools generate deterministic estimates of travel times without accounting for variability in results (due to service reliability issues, for example) ... [rafa: the idea here is to emphasize the power of R5 with the time_window parameter. We certainly need to rephrase things, but perhaps we should stress this idea here -->

@@

The r5r package has low data requirements and is easily scalable, allowing fast computation of routes and travel times at either city or country-level analysis. It creates a routable transport network using street network data from [OpenStreetMap](https://www.openstreetmap.org/) (OSM), and it automatically combines public transport data if they are provided by the user in the standard [General Transit Feed Specification](https://developers.google.com/transit/gtfs/) (GTFS) format.

@@

-   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>. This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file and stores it locally for future use; and (2) combines the OSM and GTFS datasets to build a routable network object.

@@

Both routing functions are versatile so users can easily set customized inputs such as transport modes, departure date and times, walking and cycling speeds, maximum trip duration, walking distances and number of public transport transfers. In the following section, we will focus on results obtained from `travel_time_matrix()`.

@@

After installed with the `install.packages("r5r")` command, the package can be attached (alongside some other packages to reproduce this article), as follows:

@@

### Building a routable transport network

To build a routable transport network with r5r and load it in memory, the user need to run a command line pointing to the location on disk where the input data is stored. In the examples of this paper, we will be using the provided Porto Alegre dataset.

@@

### Calculating a travel time matrix

The `travel_time_matrix()` function takes, as inputs, the spatial location of origins/destinations (either as a spatial `sf POINT` object, or as a `data.frame` containing the columns `id`, `lon` and `lat`) and a few travel parameters such as *maximum trip duration* or *walking distance*. It outputs travel time estimates for each origin-destination pair at a set `departure_datetime`.

Since service levels can significantly vary across the day [@stepniak2019impact], users can account for this variation using the `time_window` parameter. In this case, several matrices are calculated using every minute within the set time window as a departure time. By default the function outputs the 50th percentile (i.e. the median) travel time between each origin and destination pair within the time window. The `percentiles` parameter allows the user to retrieve travel time estimates at different points of the distribution. An example of the function's usage is presented below:
